---
title: "Final Exam R"
author: "Ana Frandell"
date: "March 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

```{r package-loading, echo=FALSE, message=FALSE}
library("dplyr") 
library("RSQLite") 
library("nycflights13") 
library("ggplot2")
library("magrittr")
library("reshape2")
```

```{r echo=FALSE, message=FALSE}
my_db <- nycflights13_sqlite()
flights_sqlite <- tbl(my_db, "flights")
weather_sqlite <- tbl(my_db, "weather")
airports_sqlite<- tbl(my_db, "airports")
planes_sqlite  <- tbl(my_db, "planes")
airlines_sqlite<-tbl(my_db, "airlines")
```

# Use date from NCY Flights 13

```{r echo=FALSE, warning=FALSE, message=FALSE}
flights<- data.frame(flights_sqlite, header=TRUE)
weather<- data.frame(weather_sqlite, header=TRUE)
airport <- data.frame(airports_sqlite, header=TRUE)
planes<- data.frame(planes_sqlite, header=TRUE)
airlines <-data.frame(airlines_sqlite, header=TRUE)
```

```{r echo=FALSE, message=F, warning=F}
flights$can <- ifelse(is.na(flights$arr_time), 1, 0)
#generate dummy variable for canceled flights
```

```{r echo=FALSE, message=F, warning=F}
plane.flight <- left_join((flights), (planes), by = "tailnum")
airp.flight <-  left_join(flights, airports, by = c("dest"= "faa"))
a.p.f <- left_join(airp.flight, planes, by ="tailnum")

flights <- flights %>%
  mutate(
    date = as.Date(paste(month, year, day, hour, sep = "."), format = "%m.%Y.%d.%H")
  )
weather.f <- left_join(flights, weather,
                       by=c("year", "month", "day", "hour"))
#combine, mutate, merge data
```

## Part A - Weather 

```{r echo=FALSE, message=F, warning=F}
lm(weather.f$can ~ weather.f$humid)
weather.reg <- lm(weather.f$can ~ weather.f$humid + weather.f$wind_speed + weather.f$precip + weather.f$temp)
```

```{r echo=TRUE, message=F, warning=F}
summary(weather.reg)
```

The results from this regression shows that humidity and temperature are statistically significant variables that help explain how the weather impacts canceled calls. 


```{r echo=TRUE, message=F, warning=F}
desc<-arrange(weather.f, desc(dep_delay))
head(desc$temp, 10)
```

Top 10 delayed flights by tempurature

```{r echo=TRUE, message=F, warning=F}
head(desc$wind_gust, n=10)
```

Top 10 delayed flights by wind gust 


## Part B - Aspects of Time 

```{r echo=FALSE, message=FALSE, warning=F, fig.cap = "Departure Delay by Arrival Times, ggplot"}
minute.delay<-ggplot(flights, 
             aes(x=dep_delay, y=air_time))
minute.delay+geom_point(aes(colour=origin, alpha=0.5))
```

This plot shows that less air time shows that the flight is less likely to have a departure delay. 


```{r echo=FALSE, message=F, warning=F, fig.cap = "Departure Delay by Departure Times, ggplot"}
minute.delay<-ggplot(flights, 
                     aes(x=dep_delay, y=dep_time))
minute.delay+geom_point(aes(colour=origin, alpha=0.5))
```

Depending on the origin, the flights that depart later in the day are more likely to experience departure delay.  

```{r echo=FALSE, warning=F, message=F}
flights.summary<- flights %>%
  dplyr::group_by(minute) %>%
  dplyr::summarise(
    flight.cancled.n = sum(can, na.rm= TRUE),
    flights.delay.n = sum(dep_delay, na.rm= TRUE),
    delay.min = min(dep_delay, na.rm = TRUE),
    delay.mean = mean(dep_delay, na.rm = TRUE),
    delay.median = median(dep_delay, na.rm = TRUE),
    delay.max = max(dep_delay, na.rm = TRUE)
  )
```

```{r echo=TRUE}
summary(flights.summary)
```


```{r echo=TRUE, message=F, warning=F}
hour.reg <- (lm(weather.f$can ~ weather.f$hour))
```

```{r echo=TRUE, message=F, warning=F}
summary(hour.reg)
```

This regression results shows that hours is not a statistically significant determinant for canceled flights.  

```{r echo=TRUE, message=F, warning=F}
fl.m <-flights %>%
  dplyr::group_by(year, month) %>%     
  dplyr::summarize(
    delay.min = min(dep_delay, na.rm = TRUE),
    delay.mean = mean(dep_delay, na.rm = TRUE),
    delay.max = max(dep_delay, na.rm = TRUE)
  )
fl.m$date<- as.Date(paste(fl.m$month, fl.m$year, "01", sep="."), format = "%m.%Y.%d")
ggplot(data=fl.m, aes(x=date, y=delay.mean)) + geom_line()
```

This plot shows that the average number of delayed flights increases drastically during the begining of summer and declines immediately in July.  

```{r echo=FALSE}
plot(fl.m$delay.max, type ='l', col='blue', xlab="Date", ylab="Departure Delay")
```

This plot shows the maximum amount of delayed departures over the span of 1 year. 

```{r echo=FALSE}
par(mfrow=c(2,2))
plot(fl.m$delay.mean, type = 'l', xlab="Date", ylab="Departure Delay")
plot(fl.m$delay.min, type ='l', col='red', xlab="Date", ylab="Departure Delay")
```

These graphs compare the average amount of departure delayed flights and the minimum over time.  

## Part C - Airport Destination 

```{r echo=FALSE, warning=F, message=F, fig.cap = "Average Delayed Departures, ggplot"}
org.2 <-airp.flight %>%
  dplyr::group_by(year, month, dest) %>%     #group by 3 different variables 
  dplyr::summarize(
    air_mean= mean(dep_delay, na.rm=TRUE)
  )
org.2$date<- as.Date(paste(org.2$month, org.2$year, "01", sep="."), format = "%m.%Y.%d")
ggplot(data=org.2, aes(x=date, y=air_mean, group = dest, colour = dest)) + geom_line()
```

This graphs shows where some locations experience peaks in their average number of delayed flights.  

```{r echo=FALSE, warning=F, message=F, fig.cap = "Departure Delay by Destination, ggplot"}
destination.delay<-ggplot(weather.f, 
                     aes(x=dest, y=dep_delay))
destination.delay+geom_point()
destination.delay+geom_point(aes(colour=dest, alpha=0.5))

```


THis graphs appears to show a unifrom distribution for departure delays and destination.  

```{r echo=FALSE, warning=F, message=F}
dest.can.reg <- lm(a.p.f$can ~ a.p.f$dest)
```

```{r echo=FALSE, warning=F, message=F}
summary(dest.can.reg)
```

The regression ressults shows where destination is a statistically significant variable for explaining canceled flights. 


#Part D - Characteristics of Plane

```{r echo=FALSE, warning=F, message=F}
plane.seats<-ggplot(plane.flight, 
                     aes(x=seats, y=dep_delay))
plane.seats+geom_point(aes(colour=origin, alpha=0.5))
```

This graph suggests a uniform distribution for the number of seats of a plane and the number of delayed departures.  The observations are identified by origin.  

```{r echo=FALSE, warning=F, message=F}
eng.can <- lm(plane.flight$can ~ plane.flight$engines)
summary(eng.can)
```

The regression results shows that the number of engines in the plane has a negative relationship with canceled flights.  

```{r echo=FALSE, warning=F, message=F}
eng.delay <- lm(plane.flight$dep_delay ~ plane.flight$engines)
summary(eng.delay)
```

The regression results shows that the number of engines in the plane has a positive relationship with delayed departures.  


```{r echo=FALSE, warning=F, message=F}
desc.type<-arrange(plane.flight, desc(dep_delay))
desc.flight<-arrange(flights, desc(dep_delay))
```
```{r echo=F, message=F, warning=F}
head(desc.type$seats, 10)
```

Number of seats on the top 10 delayed departure flights 

```{r echo=FALSE, warning=F, message=F}
head(desc.type$carrier, n=10)
```

Carriers of top 10 delayed departure flights.  
